{RandomRationalPoints} package for \emph{Macaulay2}

### 2021-04-20 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt

e) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.

MAIN CHANGES:

-- We have implemented the dimViaBezout to compute the dimension probabilistically. This intersects V(I) with successively higher dimensional random linear spaces until there is an intersection.  For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a 2-dimensional linear space, and so on.  This speeds up many computations.

    Some comments on implementation.  We have implemented this in the affine case and the projective case.

    In the projective case, saturate against all coordinates, instead of just the last coordinate.  Otherwise, we were getting incorrect answers too frequently.

    We also implemented this in the affine setting affine examples where we do not saturate.

    In the projective case, the referee suggested a binary search pattern, and we still use that there.   In the affine case however, we found that in practice, it is much faster to start with a 1-dimensional linear space, then a 2-dimensional linear space, and so on, until we find an intersection.  This is because in our examples, checking whether an intersection is empty is much faster when the intersection is indeed empty.

    Unfortunately, there are large classes of examples where either the probablistic code, for either the affine or non-affine examples, are much slower than simply calling dim.  We would like to implement both simultaneously by multi-threading.  Unfortunately, canceling tasks in multithreaded examples still creates a great deal of instability in M2.

@@

Additionally, we have had quite a few examples, especially when running something iteratively, where the saturation:
Js:=saturateInGenericCoordinates J;
(which only saturates in the last coordinate)
will saturate away the points, and thus causes this function to crash.

However, doing a saturation with respect to every variable (even separately) seems to make this function quite a bit slower.  We are tempted to saturate with respect to a (random) variable, then use a try command to see if that yields a good result, and if it crashed try saturating with respect to a different (random) variable.  However, we

New documentation and examples have been provided.

All the strategies now make use of the dimViaBezout to compute the dimension of the given ideal, if the codimension is not already known. In case that it is known, it should be specified along with the call of randomPoints.

### 2021-04-24 FastLinAlg/Paper/RandomRationalPoints.tex
Minor Changes on Bib File and Paper

    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (ie, with many linear or binomial terms).  Such simple linear spaces are the ones considered in {\tt randomKRationalPoint} for instance.  In practice, this seems give approximately the same performance as projecting to a hypersurface, without the chance of the code getting hung up on the generic projection.
\end{remark}

>>>>

    On the other hand, instead of using a truly random linear space to intersect with, in the default strategy, we normally choose a linear space whose defining equations are particularly non-complicated (i.e., with many linear or binomial terms).  Such simple linear spaces are the ones considered in {\tt randomKRationalPoint} for instance.  In practice, this seems give approximately the same performance as projecting to a hypersurface, without the chance of the code getting hung up on the generic projection.
\end{remark}
